---
title: "Analyzing and Modeling Motortrends Car Data Set"
author: "dortje"
date: "19 November 2017"
output:
  slidy_presentation:
    theme: null
    highlight: null
---
<style type="text/css">

body{ /* Normal  */
    font-size: 10px;
  }
h1.title {
  font-size: 38px;
  font-family: Helvetica;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 38px;
  font-family: Helvetica;
  color: DarkRed;
}
h2 { /* Header 2 */
    font-size: 32px;
    font-family: Helvetica;
    color: DarkRed;
}
h3 { /* Header 3 */
  font-size: 22px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
```
## Introduction
With the super awesome shiny app "Analyzing and Modeling Motortrends Car Data Set" you never need to worry about the mtcars data set again^1^. The app helps with two things:

- exploring the variables within the data set while also paying attention to the relationship and
- modling a linear regression model by choosing the outcome and then selecting the predictors.



^1^ When it comes to building linear regression models on the data set.

## Analyzing
Assuming the user selected "Horse Power" for the x axis, "Miles per Gallon" for the y axis and chose "Number of Cylinders" as the category:
```{r analyzing, echo=TRUE, eval=TRUE}
x_var <- "hp"; y_var <- "mpg"; col_var <- "cyl"
g <- ggplot(mtcars, aes_string(x=x_var, y=y_var, color=col_var))
g + geom_point() + geom_smooth(method="lm")
```

## Modeling
Assuming the user selected "Miles per Gallon" as output and "Dispension" along with "Cylinder" as predictors:
```{r modeling, echo=TRUE, eval=TRUE}
out <- "mpg"; preds <- c("disp","cyl")
model_formula = formula(paste(out, "~", paste(preds, collapse="+")))
lin_model <- lm(model_formula, data=mtcars)
summary(lin_model)
```

## Final Notes

Please check out the awesome app and tell me what you think about it:
https://dortje.shinyapps.io/mtcars_visualization/

Some Background Information:

- The app is robust to some errors (e.g. selecting no predictors or selecting the output also as a predictor).
- To make the visualizations dynamic, the input strings need to be transformed to variables with the help of aes_string() for ggplot() and formula() for creating the formula for the linear model.