---
title: "Analyzing and Modeling Motortrends Car Data Set"
author: "dortje"
date: "19 November 2017"
output:
  slidy_presentation:
    theme: null
    highlight: null
---
<style type="text/css">

body{ /* Normal  */
    font-size: 10px;
  }
h1.title {
  font-size: 38px;
  font-family: Helvetica;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 38px;
  font-family: Helvetica;
  color: DarkRed;
}
h2 { /* Header 2 */
    font-size: 32px;
    font-family: Helvetica;
    color: DarkRed;
}
h3 { /* Header 3 */
  font-size: 22px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
```
## Introduction
With the super awesome shiny app "Analyzing and Modeling Motortrends Car Data Set" you never need to worry about the mtcars data set again^1^. The app helps with two things:

- exploring the variables within the data set while also paying attention to the relationship and
- modling a linear regression model by choosing the outcome and then selecting the predictors.

Please check out the awesome app and tell me what you think about it:
https://dortje.shinyapps.io/mtcars_visualization/

^1^ When it comes to building linear regression models on the data set.

## Analyzing
Assuming the user selected "Horse Power" for the x axis, "Miles per Gallon" for the y axis and chose "Number of Cylinders" as the category:
```{r analyzing, echo=TRUE, eval=TRUE}
x_var <- "hp"; y_var <- "mpg"; col_var <- "cyl"
g <- ggplot(mtcars, aes_string(x=x_var, y=y_var, color=col_var))
g + geom_point() + geom_smooth(method="lm")
```

## Modeling
Assuming the user selected "Miles per Gallon" as output and "Dispension" along with "Cylinder" as predictors:
```{r modeling, echo=TRUE, eval=TRUE}
out <- "mpg"; preds <- c("disp","cyl")
model_formula = formula(paste(out, "~", paste(preds, collapse="+")))
lin_model <- lm(model_formula, data=mtcars)
summary(lin_model)
```

## Full code from server.R
```{r server_code, echo=TRUE, eval=FALSE}
library(shiny)
library(ggplot2)

data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- as.factor(mtcars$am)
mtcars$gear <- as.factor(mtcars$gear)
mtcars$carb <- as.factor(mtcars$carb)

shinyServer(function(input, output) {
    lin_model <- reactive({
        out <- input$select_out
        preds <- input$chkbox_pred
        preds <- preds[!preds %in% out]
        
        if(length(preds) > 0) {
                model_formula = formula(paste(out, "~", paste(preds, collapse="+")))
                lm(model_formula, data=mtcars)
        } else {
           return(NULL) 
        }
    })
    output$var_plot <- renderPlot({
        x_var <- input$select_x
        y_var <- input$select_y
        col_var <- input$select_col
        if (col_var == "none") {
            g <- ggplot(mtcars, aes_string(x=x_var, y=y_var))
        } else {
            g <- ggplot(mtcars, aes_string(x=x_var, y=y_var, color=col_var))
        }
        g + geom_point() + geom_smooth(method="lm")

    })
    output$modelSummary <- renderPrint({
        if(!is.null(lin_model())) {
            summary(lin_model())
        } else {
            print("No predictors selected yet.")
        }
    })
})
```